### 2019-01-15 R/pdf analysis of pages in wg reports.Rmd

---
output: 
  word_document:
    reference_docx: ../report_template_v1.5.dotx
---

```{r, message=FALSE, warning=FALSE, include=FALSE}

# -----------------------------------------------------------------------------------------------
# PDF analysis of pages in WG reports.Rmd
#
# 11/01/2019 first coding; adapted from ICES advice and assessment summaries.Rmd
# -----------------------------------------------------------------------------------------------

require("knitr")
knitr::opts_chunk$set(echo = FALSE,	message = FALSE,	warning = FALSE,	comment = "",	crop = TRUE )
knitr::opts_chunk$set(fig.width=10) 
knit_hooks$set(crop = hook_pdfcrop)

# Libraries
library(pdftools)
library(tidyverse)

# captioner library
library(captioner) # deal with captions
fig_nums <- captioner::captioner(prefix = "Figure ", levels=1, type=c("n"))

# Load utils code
source("../../mptools/r/my_utils.r")

# file path
my.path <- "D:/ZZZ"

# counter for loop
i <- 0

# loop over years
for (wg in c("HAWG", "WGNSSK", "AFWG", "WGBFAS", "WGCSE", "WGSSDS", "WGNSDS", "WGWIDE", "WGMHSA","WGDEMPELBAL", 
             "WGHMM", "WGBIE", "WGDEEP", "WGANSA", "WGHANSA", "WGEF", "NWWG", "WGNEW")) {

  for (y in 2000:2018) {

    fn <- file.path(my.path, paste0(wg," ",y,".pdf")) 
    # print(fn)
    if (file.exists(fn)) {
      # print(pdf_info(fn)[2])
      t <- data.frame(wg=wg, year=y, npages = pdf_info(fn)[2], stringsAsFactors = FALSE)
      if (i == 0) {wg_pages <- t} else {wg_pages <- bind_rows(wg_pages, t)}
      i <- i + 1
    }
  } # end of y loop
}  # end of wg loop

# print(wg_pages)

# summarize by pentad of 5 years
wg_pages_bypentad <-
  wg_pages %>% 
  group_by(year) %>% 
  summarize(pages = sum(pages)) %>% 
  mutate(
    pentad = 10 / 2 * floor((year*2)/10)
  ) %>% 
  group_by(pentad) %>% 
  mutate(avg = mean(pages),
         end = pentad + 4)

# Figure numbers
fig_nums(name    = "pages_bywg", level = 1, display = FALSE,
         caption = "Number of pages by WG")

fig_nums(name    = "pages_byyear", level = 1, display = FALSE,
         caption = "Number of pages by year and averages by pentad")

```

**Number of pages in WG reports**

M.A. Pastoors

`r format(Sys.time(), '%d/%m/%Y')`

Number of pages in WG reports of the following ICES working groups: `r unique(wg_pages$wg)`.

```{r, echo=FALSE, fig.asp=0.7, fig.align="center", message=FALSE, warning=FALSE}

wg_pages %>%

  ggplot(aes(x=year,y=pages)) +
  theme_publication() +
  theme(legend.position = "none") +
  geom_bar(aes(fill=wg), stat="identity", colour="white", size=0.2) +
  expand_limits(y=0) +
  scale_x_continuous(breaks=scales::pretty_breaks()) +
  facet_wrap(~wg)

```

*`r fig_nums("pages_bywg")`*

```{r, echo=FALSE, fig.asp=0.7, fig.align="center", message=FALSE, warning=FALSE}

wg_pages %>%

  ggplot(aes(x=year,y=pages)) +
  theme_publication() +
  theme(legend.title = element_blank()) +
  geom_bar(aes(fill=wg), stat="identity", colour="white", size=0.2) +
  geom_segment(data=wg_pages_bypentad, aes(x=pentad, xend=end, y=avg,yend=avg), colour="black", size=2) +
  geom_text   (data=wg_pages_bypentad, aes(x=pentad, y=avg, label=as.integer(avg)), 
               vjust=-0.5, hjust=0) +
  expand_limits(y=0) +
  scale_x_continuous(breaks=scales::pretty_breaks())

```

*`r fig_nums("pages_byyear")`*