### 2019-09-05 R/ICES advice and assessment summaries WGWIDE 2019.Rmd
Updates after WGWIDE 2019

---
output: 
  word_document:
    reference_docx: ../report_template_v1.5.dotx
---

```{r, message=FALSE, warning=FALSE, include=FALSE}

# -----------------------------------------------------------------------------------------------
# ICES advice and assessment summaries.Rmd
#
# 06/07/2018 first coding
# 14/12/2018 updated for the new database structure
# 01/06/2019 updated for herring advice round
# 03/09/2019 updated after WGWIDE
# -----------------------------------------------------------------------------------------------

require("knitr")
knitr::opts_chunk$set(echo = FALSE,	message = FALSE,	warning = FALSE,	comment = "",	crop = TRUE )
knitr::opts_chunk$set(fig.width=10) 
knit_hooks$set(crop = hook_pdfcrop)

# Libraries
library(rmarkdown) # creating markdown documents
library(tidyverse) # combined package of dplyr, tidyr, ggplot, readr, purrr and tibble
library(reshape2)  # reshaping data; e.g. dcast
library(ggthemes)  # for themes
library(pander)    # for print tables
library(readxl)    # read excel files
library(cowplot)   # multiplots
library(directlabels)  # label graphs
library(RColorBrewer) # colours
library(lubridate)

# Load utils code
source("../../mptools/r/my_utils.r")

dropboxdir <- paste(get_dropbox(), "/iAdvice", sep="")

# Load dataset
load(file=paste(dropboxdir, "/rdata/iAssess.RData",sep=""))
load(file=paste(dropboxdir, "/rdata/iAdvice.RData",sep=""))

fay <- 2011
fy  <- 2000
lay <- 2019
ly  <- 2019

stock="hom-nsea"
assessyear=lay
assessyear=2017

firstyear=fy
lastyear=ly
plot = c("stocksize", "fishingpressure", "recruitment")
my.purpose="advice"

# Function to plot single assessment
plot_assess <- function(stock,assessyear, firstyear, lastyear, 
                        my.purpose = "advice", plot=c("stocksize","fishingpressure")) {

  d <-
    iAssess %>% 
    ungroup() %>%

    filter(stockkeylabelold %in% stock | stockkeylabel %in% stock) %>% 
    filter(assessmentyear   == assessyear) %>% 
    filter(year             %in% firstyear:lastyear) %>% 
    filter(purpose %in% my.purpose) %>%

    #set type
    mutate(
      type = ifelse(purpose == "advice"      , "last"    , NA),
      type = ifelse(grepl("bench",purpose)   ,  "benchmark", type),
      type = ifelse(grepl("replaced",purpose), "replaced"  , type),
      type = ifelse(is.na(type), "other", type)) %>%

    mutate(tyear     = substr(as.character(assessmentyear), 3,4)) %>% 
    dplyr::select(stockkeylabel, stockkeylabelold, stockdescription, type, 
                  assessmentyear, year,
                  recruitment, lowrecruitment, highrecruitment, unitofrecruitment,
                  stocksize, lowstocksize, highstocksize, stocksizeunits,
                  fishingpressure, lowfishingpressure, highfishingpressure, fishingpressureunits,
                  landings, discards, catches, catcheslandingsunits) %>%

    # convert to long data
    gather(key=variable, value=data, c(recruitment, lowrecruitment, highrecruitment, 
                  stocksize, lowstocksize, highstocksize,
                  fishingpressure, lowfishingpressure, highfishingpressure, 
                  landings, discards, catches)) %>%

    # set appropriate units and remove old units columns
    mutate(
      unit = ifelse(grepl("recruitment", variable), unitofrecruitment, NA),
      unit = ifelse(grepl("stocksize"  , variable), stocksizeunits, unit),
      unit = ifelse(grepl("fishing"    , variable), fishingpressureunits, unit),
      unit = ifelse(grepl("landing|discard|catch", variable), catcheslandingsunits, unit) ) %>% 
    dplyr::select(-unitofrecruitment, - stocksizeunits, -fishingpressureunits, -catcheslandingsunits) %>%

    # separate estimate, low and high
    mutate(
      esttype = ifelse(grepl("low", variable), "low", NA),
      esttype = ifelse(grepl("high", variable), "high", esttype),
      esttype = ifelse(is.na(esttype), "est", esttype) ) %>% 
    mutate(
      variable = gsub("low|high","", variable) ) %>%

    # select variables to plot
    filter(variable %in% plot) %>% 
    mutate(variable = factor(variable, levels=plot, ordered=TRUE)) %>%

    # spread to make separate columns for est, low and high
    spread(key=esttype, value=data)

  # Reference points
  rp <-
    iAdvice %>% 
    ungroup() %>% 
    filter(purpose %in% my.purpose) %>% 
    filter(stockkeylabelold %in% stock | stockkeylabel %in% stock) %>% 
    filter(assessmentyear   == assessyear) %>% 
    dplyr::select(stockkeylabel, stockkeylabelold, assessmentyear,
              blim, bpa, msybtrigger, flim, fpa, fmsy) %>%
    gather(key=refpoint, value=data, blim:fmsy) %>%

    # generate variables
    mutate(
      variable = ifelse(refpoint %in% c("blim","bpa","msybtrigger"), "stocksize", NA),
      variable = ifelse(refpoint %in% c("flim", "fpa", "fmsy"), "fishingpressure", variable) ) %>% 
    mutate(variable = factor(variable, levels=plot, ordered=TRUE)) %>%

    # generate refpoint types
    mutate(
      reftype = ifelse(refpoint %in% c("blim","flim"), "lim", NA),
      reftype = ifelse(refpoint %in% c("bpa", "fpa") , "pa", reftype),
      reftype = ifelse(refpoint %in% c("msybtrigger", "fmsy"), "msy", reftype))  %>% 
    dplyr::select(-refpoint)

  # Titles
  tt <-
    d %>% 
    distinct(variable, unit, stockkeylabel, stockdescription) %>%
    filter(!is.na(unit))

  # d %>% View()

  # plot 
  d %>%

    ggplot(aes(x=year,y=est)) +

    theme_publication() +

    theme(legend.title=element_blank(),
          axis.text.x = element_text(angle = 0, vjust = 0.5, size=9),
          axis.text.y = element_text(size=9),
          plot.title  = element_text(size=10, hjust=0), 
          plot.margin = margin(2, 2, 2, 2, "mm"),
          # strip.background = element_blank(),
          legend.position = "null") +

    geom_ribbon(aes(x=year, ymin=low, ymax=high, fill=type), alpha=0.2) +

    geom_line(aes(colour = type, size=type, linetype=type) ) +

    # geom_ribbon(aes(ymin=low, ymax=high), alpha=0.2, fill="red") +
    # geom_line(colour="red" ) +

    # plot limits
    geom_hline(data=filter(rp, reftype %in% "lim"), aes(yintercept=data), 
               colour="red", linetype="solid", inherit.aes=FALSE) +
    geom_text (data=filter(rp, reftype %in% "lim"), aes(x=firstyear, y=data), 
               label = "lim", vjust = -0.3, hjust = 0, colour="red", inherit.aes=FALSE) +

    # plot msy
    geom_hline(data=filter(rp, reftype %in% "msy"), aes(yintercept=data), 
               colour="red", linetype="dashed", inherit.aes=FALSE) +
    geom_text (data=filter(rp, reftype %in% "msy"), aes(x=firstyear, y=data), 
               label = "msy", vjust = -0.3, hjust = 0, colour="red", inherit.aes=FALSE) +

    scale_colour_manual  (
      values=c(last ="red",last_bench="blue",advice="black",
               benchmark="blue",replaced="green",other="gray80")) +
    scale_fill_manual    (
      values=c(last = "red",last_bench="blue",advice="black",
               benchmark="blue",replaced="green",other="gray80")) +
    scale_linetype_manual(
      values=c(last="solid",last_bench="solid",advice="solid",
               benchmark="solid",replaced="solid",other="solid")) +
    scale_size_manual    (
      values=c(last= 1.2,last_bench=1.2,advice=0.8,
               benchmark=0.8,replaced=0.8,other=0.8)) +

    expand_limits(y = 0) +
    xlim(firstyear,lastyear) +
    labs(x = NULL, y = NULL, 
         title = paste0(unique(tt$stockdescription), " (", unique(tt$stockkeylabel),")") ) +
    facet_wrap(~variable, scales="free_y")

  } # End of plot_assess function

# s   <- "mac-nea"
# plot_assess(stock=s,assessyear=lay, firstyear=fy, lastyear=ly, 
#             plot = c("stocksize", "fishingpressure", "recruitment"), 
#             my.purpose="advice")
# plot_assess(stock=s,assessyear=lay, firstyear=fy, lastyear=ly, 
#             plot = c("stocksize", "fishingpressure", "recruitment"), 
#             my.purpose="bench")

stock="whb-comb"
firstassessyear=fay; lastassessyear=lay; firstyear=fy;lastyear=ly;
plot = c("stocksize", "fishingpressure")
include.benchmark=TRUE; include.replaced=TRUE
plot.uncertainty=FALSE

# Function to plot historical assessments
plot_history <- function(stock,firstassessyear,lastassessyear, firstyear, lastyear, 
                         include.benchmark = TRUE, include.replaced=TRUE, 
                         plot=c("stocksize","fishingpressure"),
                         plot.uncertainty=FALSE) {

  d <-
    iAssess %>% 
    ungroup() %>%

    {if (include.benchmark == TRUE & include.replaced == TRUE) {
      filter(., purpose %in% c("advice","benchmark", "interbenchmark", "bench", "replaced"))
    } else if (include.benchmark == TRUE & include.replaced == FALSE) {
      filter(., purpose %in% c("advice","benchmark", "interbenchmark", "bench"))
    } else if (include.benchmark == FALSE & include.replaced == TRUE) {
      filter(., purpose %in% c("advice","replaced"))
    } else {
      filter(., purpose %in% c("advice")) 
    }} %>%

    filter(stockkeylabelold %in% stock | stockkeylabel %in% stock) %>% 
    filter(assessmentyear   %in% c(firstassessyear:lastassessyear)) %>% 
    filter(year             %in% firstyear:lastyear) %>%

    group_by(assessmentyear) %>%

    #set type
    mutate(
      type = ifelse(assessmentyear == lastassessyear & purpose == "advice"      , "last"      , NA),
      type = ifelse(assessmentyear == lastassessyear & grepl("bench",purpose) & 
                    n_distinct(purpose)==1, "last_bench", type),
      type = ifelse(assessmentyear == lastassessyear & grepl("bench",purpose) & 
                      n_distinct(purpose) > 1, "bench", type),
      type = ifelse(assessmentyear != lastassessyear & purpose == "advice"      , "advice"    , type),
      type = ifelse(assessmentyear != lastassessyear & grepl("bench",purpose)   ,  "benchmark", type),
      type = ifelse(assessmentyear != lastassessyear & grepl("replaced",purpose), "replaced"  , type),
      type = ifelse(is.na(type), "other", type)) %>%

    ungroup() %>%

    # set tyear and group
    mutate(tyear = substr(as.character(assessmentyear), 3,4)) %>% 
    mutate(group = paste0(tyear, type)) %>% 
    mutate(group = substr(group, 1,3)) %>% 
    mutate(group = gsub("a|l","", group)) %>%

    dplyr::select(stockkeylabel, stockkeylabelold, stockdescription, 
                  assessmentyear, year, tyear, group, type, purpose,
                  recruitment, lowrecruitment, highrecruitment, unitofrecruitment,
                  stocksize, lowstocksize, highstocksize, stocksizeunits,
                  fishingpressure, lowfishingpressure, highfishingpressure, fishingpressureunits,
                  landings, discards, catches, catcheslandingsunits) %>%

    # convert to long data
    gather(key=variable, value=data, 
           c(recruitment, lowrecruitment, highrecruitment, 
             stocksize, lowstocksize, highstocksize,
             fishingpressure, lowfishingpressure, highfishingpressure, 
             landings, discards, catches)) %>%

    # set appropriate units and remove old units columns
    mutate(
      unit = ifelse(grepl("recruitment", variable), unitofrecruitment, NA),
      unit = ifelse(grepl("stocksize"  , variable), stocksizeunits, unit),
      unit = ifelse(grepl("fishing"    , variable), fishingpressureunits, unit),
      unit = ifelse(grepl("landing|discard|catch", variable), catcheslandingsunits, unit) ) %>% 
    dplyr::select(-unitofrecruitment, - stocksizeunits, -fishingpressureunits, -catcheslandingsunits) %>%

    # separate estimate, low and high
    mutate(
      esttype = ifelse(grepl("low", variable), "low", NA),
      esttype = ifelse(grepl("high", variable), "high", esttype),
      esttype = ifelse(is.na(esttype), "est", esttype) ) %>% 
    mutate(
      variable = gsub("low|high","", variable) ) %>%

    # select variables to plot
    filter(variable %in% plot) %>% 
    mutate(variable = factor(variable, levels=plot, ordered=TRUE)) %>%

    # spread to make separate columns for est, low and high
    spread(key=esttype, value=data) %>%

    # remove NA
    filter(!is.na(est))

  # Check
  # d %>% filter(assessmentyear == 2019) %>% View()

  # Reference points
  rp <-
    iAdvice %>% 
    ungroup() %>% 
    filter(purpose %in% my.purpose) %>% 
    filter(stockkeylabelold %in% stock | stockkeylabel %in% stock) %>% 
    filter(assessmentyear   == assessyear) %>% 
    dplyr::select(stockkeylabel, stockkeylabelold, assessmentyear,
              blim, bpa, msybtrigger, flim, fpa, fmsy) %>%
    gather(key=refpoint, value=data, blim:fmsy) %>%

    # generate variables
    mutate(
      variable = ifelse(refpoint %in% c("blim","bpa","msybtrigger"), "stocksize", NA),
      variable = ifelse(refpoint %in% c("flim", "fpa", "fmsy"), "fishingpressure", variable) ) %>% 
    mutate(variable = factor(variable, levels=plot, ordered=TRUE)) %>%

    # generate refpoint types
    mutate(
      reftype = ifelse(refpoint %in% c("blim","flim"), "lim", NA),
      reftype = ifelse(refpoint %in% c("bpa", "fpa") , "pa", reftype),
      reftype = ifelse(refpoint %in% c("msybtrigger", "fmsy"), "msy", reftype))  %>% 
    dplyr::select(-refpoint)

  # Titles
  tt <-
    d %>% 
    distinct(variable, unit, stockkeylabel, stockdescription) %>%
    filter(!is.na(unit))

  # plot
  d %>%

    ggplot(aes(year,est, group=group)) +

    theme_publication() +

    theme(legend.title=element_blank(),
          axis.text.x = element_text(angle = 0, vjust = 0.5, size=9),
          axis.text.y = element_text(size=9),
          plot.title  = element_text(size=10, hjust=0), 
          plot.margin = margin(2, 2, 2, 2, "mm"),
          # strip.background = element_blank(),
          legend.position = "null") +

    {if (plot.uncertainty)     
      geom_ribbon(data=filter(d, type=="last"),
          aes(x=year, ymin=low, ymax=high, fill=type), alpha=0.2) } +

    geom_line(aes(colour = type, size=type, linetype=type) ) +

    geom_dl(aes(label  = group, colour = type, group=group), 
            method = list(dl.combine("last.points"), cex = 0.8, inherit.aes=FALSE)) +

    # plot limits
    geom_hline(data=filter(rp, reftype %in% "lim"), aes(yintercept=data), 
               colour="red", linetype="solid", inherit.aes=FALSE) +
    geom_text (data=filter(rp, reftype %in% "lim"), aes(x=firstyear, y=data), 
               label = "lim", vjust = -0.3, hjust = 0, colour="red", inherit.aes=FALSE) +

    # plot msy
    geom_hline(data=filter(rp, reftype %in% "msy"), aes(yintercept=data), 
               colour="red", linetype="dashed", inherit.aes=FALSE) +
    geom_text (data=filter(rp, reftype %in% "msy"), aes(x=firstyear, y=data), 
               label = "msy", vjust = -0.3, hjust = 0, colour="red", inherit.aes=FALSE) +

    scale_colour_manual  (
      values=c(last ="red",last_bench="blue",advice="black",
               benchmark="blue",replaced="green",other="gray80")) +
    scale_fill_manual    (
      values=c(last = "red",last_bench="blue",advice="black",
               benchmark="blue",replaced="green",other="gray80")) +
    scale_linetype_manual(
      values=c(last="solid",last_bench="solid",advice="solid",
               benchmark="solid",replaced="solid",other="solid")) +
    scale_size_manual    (
      values=c(last= 1.2,last_bench=1.2,advice=0.8,
               benchmark=0.8,replaced=0.8,other=0.8)) +

    expand_limits(y = 0) +
    xlim(firstyear,lastyear) +
    labs(x = NULL, y = NULL, 
         title = paste0(unique(tt$stockdescription), " (", unique(tt$stockkeylabel),")") ) +
    facet_wrap(~variable, scales="free_y")

  # return(d)

  } # End of plot_history function

# firstadviceyear <- 2016; lastadviceyear  <- 2020; stock<-c("mac-nea"); include.replaced <- TRUE

# Function to plot advice overviews
table_advice <- function(stock,firstadviceyear,lastadviceyear, include.replaced=FALSE) {

  d <-
    iAdvice %>%

    filter(stockkeylabelold %in% stock | stockkeylabel %in% stock) %>%

    filter(tacyear %in% firstadviceyear:lastadviceyear) %>%

    {if (include.replaced == TRUE) {
      filter(., purpose %in% c("advice","replaced"))
    } else {
      filter(., purpose %in% c("advice")) 
    }} %>%

    mutate(advicebasis= substr(advicebasis,1,25),

           advice     = ifelse(!is.na(advisedcatchmax), advisedcatchmax, advisedlandingsmax),
           advicemax  = ifelse(!is.na(advisedcatchmax), advisedcatchmax, advisedlandingsmax),
           advicetype = ifelse(!is.na(advisedcatchmax), "catch", "landings"),

           tac        = ifelse(!is.na(tac), tac, tal),
           tactype    = ifelse(!is.na(tac), "catch", "landings"),

           catch      = ifelse(!is.na(catches), catches, NA),
           catchtype  = ifelse(!is.na(catches), "catches", ""),

           catchtype  = ifelse(!is.na(landings) & is.na(catch), "landings", catchtype),
           catch      = ifelse(!is.na(landings) & is.na(catch), landings, catch),

           catchtype  = ifelse(!is.na(officiallandings) & is.na(catch), "officiallandings", catchtype),
           catch      = ifelse(!is.na(officiallandings) & is.na(catch), officiallandings, catch),

           adv_c      = 100 * (advicemax / lag(advicemax, n=1) - 1),
           tac_c      = 100 * (tac /lag(tac, n=1) - 1),
           cat_c       = 100 * (catch / lag(catch, n=1) -1)
           ) %>%

    select(stockkeylabel, tacyear, advicebasis, purpose, advice, fadvmax, advicemax, tac, catch, 
           adv_c, tac_c, cat_c,
           advicetype, tactype, catchtype)

  d %>% 
  select("stockkeylabel", "tacyear", "advicebasis", "purpose", "fadvmax", "advicemax", "tac", "catch") %>%
  arrange(tacyear, desc(purpose)) %>% 
  pandoc.table(., 
               style        = "simple",
               split.tables = 200, 
               justify      = "right",
               missing      =" ",
               big.mark     = '', 
               round        = c(0,0,0,2,2,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0))
}

# table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1, include.replaced = TRUE)

```

**Summary of ICES 2019 WGWIDE assessments and advice**

Martin Pastoors, `r format(Sys.time(), '%d/%m/%Y')`

In this document an overview is presented of the ICES assessment and advice that resulted from the ICES Working Group on Widely Distributed Stocks (WGWIDE 2019) which deals with mackerel, blue whiting, horse mackerel and Atlanto-scandian herring. The overviews are based on two databases that have been developed as part of a more comprehensive research effort into the role of fisheries advice in European decision-making: one database containing the history of stock assessments and one database containing the history of scientific advice and management decisions.

**Northeast Atlantic Mackerel**

The ICES advice on Northeast Atlantic mackerel that was released in October 2018 has been received with a lot of scepticism and unbelief. According to that advice, the stock had been declining since 2011 and was estimated to be just below the MSY Btrigger. At the time, it was already found that the model results were very sensitive to the inclusion of the RFID tags. Leaving out only the RFID tags, gave a much higher stock estimate. To deal with this assessment problem, an inter-benchmark was set up which is lead to a new assessment configuration in March 2019 and an updated advice for 2019.

The new assessment that was carried out in WGWIDE 2019 has not created a different perception on stock trends of mackerel. The preliminary results of the triennial mackerel egg survey indicated a low SSB for mackerel. However, the other information sources indicate an increase in the recruitment and biomass. The egg survey carried a rather limited weight in the final model and therefore the assessment estimates a high stock in 2020 and fishing mortality only just above Fmsy.

```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}

s   <- "mac-nea"

plot_assess(stock=s,assessyear=lay, firstyear=fy, lastyear=ly, 
            plot = c("stocksize", "fishingpressure", "recruitment"),
            my.purpose="advice")

```

_Figure 1a Northeast Atlantic mackerel. Summary of most recent assessment_

#####page break

```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}

plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly, 
             include.benchmark = TRUE, 
             plot = c("stocksize", "fishingpressure")) 
```

_Figure 1b Northeast Atlantic mackerel. History of assessments. Red is most recent assessment, black is earlier assessments, blue is benchmarks and green is replaced assessments. _

```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}

table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1, include.replaced = FALSE)

# [mac.27.nea.pdf](http://ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/mac.27.nea.pdf)

```

_Table 1 Northeast Atlantic mackerel. History of advice, tac and catch._

#####page break

*Blue whiting*

The blue whiting survey in 2019 went well and showed the highest biomass of the time series. Despite the high catches, the stock is still at a very high level. The stock size of blue whiting is estimated to be around 5 million tonnes. The stock size of blue whiting is estimated higher than last year, due to a high survey result for the strong 2014 year class. However, the stock is currently decreasing due to low recruitments coming in. Catches have been well above 1 Million tonnes over the last few years and above the scientific advice. Fishing mortality is estimated to be above Fmsy.

The advice for 2020 is 1.16 million tonnes, which is at the same levels as the advice for 2019. A decline in catching opportunities in the future years can be expected.

```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}

s   <- "whb-comb"

plot_assess(stock=s,assessyear=lay, firstyear=fy, lastyear=ly, 
            plot = c("stocksize", "fishingpressure", "recruitment"),
            my.purpose="advice")

```

_Figure 2a Blue whiting. Summary of most recent assessment_

#####page break

```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}

s   <- "whb-comb"

plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly, 
             include.benchmark = TRUE, include.replaced = FALSE, 
             plot = c("stocksize", "fishingpressure"), 
             plot.uncertainty=FALSE)

```

_Figure 2b Blue whiting. History of assessments. Red is most recent assessment, black is earlier assessments and blue is benchmarks. _

```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}

s   <- "whb-comb"

table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1, include.replaced = FALSE)

# [whb.27.1-91214.pdf](http://ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/whb.27.1-91214.pdf)

```

_Table 2 Blue whiting. History of advice, tac and catch._

##### Page break

*Western horse mackerel*

The major discussion about Western horse mackerel was focussed on the reference points. An Interbenchmark protocol has been followed to re-estimate the reference points that had been set during the previous benchmark in 2017. At that time the Blim had been set at Bloss which was the last point of the time-series. Due to the retrospective revisions of the assessments since the benchmark, it became clear that that choice of Blim was inappropriate. The new reference points that have been proposed revert back to the earlier method of setting reference points for this stock: the Bpa / MSY Btrigger was selected as the lowest estimated stocks size in the previous low period of the stock (2003) during which no impaired recruitment was observed. The Blim and F reference points where then derived from that. The following reference points have been proposed: Blim = 834480 t (was 661917 t), Bpa = 1168272 t (was 911587 t), Fmsy = 0.074 (was 0.108).

The stock assessment of Western horse mackerel was reasonably comparable to last year's assessment. The retrospective revision is less than during the previous years. The stock is estimated just above the new Blim and fishing mortality is close to Fmsy. The recent recruitments appear to have been somewhat higher which leads to an modest increase in stock size.

An alternative assessment approach for Western horse mackerel has been presented in WGWIDE. This is using the SAM assessment model and largely the same data as in the standard assessment model (Stock Synthesis). The trends in the stock and fishing mortality are largely comparable although the absolute levels and uncertainties are different (higher F, lower stock and higher uncertainty in the SAM assessment).

The advice is based on the new reference points. Because the stock is between MSY Btrigger and Blim, a lower fishing mortality than Fmsy is used as the basis of the advice, which results in a substantially lower advice compared to last year. Given the indications we get from the fisheries, this could be a more realistic advice than the advice that was presented last year.

```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}

s   <- "hom-west"

plot_assess(stock=s,assessyear=lay, firstyear=fy, lastyear=ly, 
            plot = c("stocksize", "fishingpressure", "recruitment"),
            my.purpose="advice")

```

_Figure 3a Western horse mackerel. Summary of most recent assessment_

##### page break

```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}

s   <- "hom-west"

plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly, 
             include.benchmark = TRUE, include.replaced = FALSE, 
             plot = c("stocksize", "fishingpressure"), 
             plot.uncertainty=FALSE)

```

_Figure 3b Western horse mackerel. History of assessments. Red is most recent assessment, black is earlier assessments and blue is benchmarks. _

```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}

s   <- "hom-west"

table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1, include.replaced = FALSE)

# [hom.27.2a4a5b6a7a-ce-k8.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/hom.27.2a4a5b6a7a-ce-k8.pdf)

```

_Table 3 Western horse mackerel. History of advice, tac and catch._

##### page break

*North Sea horse mackerel*

The assessment and advice of North Sea is carried out every second year. A new stock assessor was responsible for the assessment (as the previous stock assessor changed jobs), and she found a number of errors in the codes that had been used in 2017. These errors were now corrected and lead to a much lower estimate of stock size in 2016.

This is a so-called category 3 advice, which means that it only uses the relative trend in a (combined) survey to derive the advice. The trend is based on the abundance of the North Sea IBTS survey (quarter 3) and French Channel CGFS survey (quarter 4). The stock is considered to be at a low level even though recruitment appears to have been higher in recent year. A length-based proxy indicator suggests that fishing mortality is probably (just) above Fmsy.

The advice is lower than the advice presented in 2017. Note that the advice is a catch advice. The advice for the landed part is 13305 t.The advice is applicable for 2020 and 2021.

```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}

s   <- "hom-nsea"

plot_assess(stock=s,assessyear=lay, firstyear=fy, lastyear=ly, 
            plot = c("stocksize", "fishingpressure", "recruitment"),
            my.purpose="advice")

```

_Figure 4a North Sea horse mackerel. Summary of most recent assessment_

##### page break

```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}

s   <- "hom-nsea"

plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly, 
             include.benchmark = TRUE, include.replaced = FALSE, 
             plot = c("stocksize", "fishingpressure"), 
             plot.uncertainty=FALSE)

```

_Figure 4b North Sea horse mackerel. History of assessments. Red is most recent assessment, black is earlier assessments and blue is benchmarks. _

```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}

s   <- "hom-nsea"

table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1, include.replaced = FALSE)

```

_Table 4 North Sea horse mackerel. History of advice, tac and catch._

##### page break

*Atlanto-scandian herring*

The stock assessment is consistent with last year's assessment. The stock appears to be slowly declining since 2009. Fishing mortality is estimated to be below Fmsy. There are signs of a reasonable sized incoming year class 2016, but this year class is not showing up in the assessment yet, as the year class is still only seen in the Barents Sea. The year class would be expected to recruit to the fishery in 2020.

The advice is lower than last year due to the lower stock size.

```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}

s   <- "her-noss"

plot_assess(stock=s,assessyear=lay, firstyear=fy, lastyear=ly, 
            plot = c("stocksize", "fishingpressure", "recruitment"),
            my.purpose="advice")

```

_Figure 5a Atlanto-scandian herring. Summary of most recent assessment_

##### page break

```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}

plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly, 
             include.benchmark = TRUE, include.replaced = FALSE, 
             plot = c("stocksize", "fishingpressure"), 
             plot.uncertainty=FALSE)

```

_Figure 5b Atlanto-scandian herring. History of assessments. Red is most recent assessment, black is earlier assessments and blue is benchmarks. _

```{r, echo=FALSE, fig.asp=0.45, fig.align="center", message=FALSE, warning=FALSE}

table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1, include.replaced = FALSE)

# [her.27.1-24a514a.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/her.27.1-24a514a.pdf)

```

_Table 5 Atlanto-scandian herring. History of advice, tac and catch._

*More information*

Please refer to the [ICES website](http://www.ices.dk/community/advisory-process/Pages/Latest-advice.aspx) for an overview of all advice documents.