### 2022-01-20 R/Presentation for pelagic forum 2019.Rmd
Pelagic forum 2019

---
output:
  powerpoint_presentation:
    reference_doc: ../powerpoint_template.pptx
---

```{r, message=FALSE, warning=FALSE, include=FALSE}

# ---
# output:
#   powerpoint_presentation:
#     reference_doc: ../powerpoint_template.pptx
# ---

# -----------------------------------------------------------------------------------------------
# Presentation for pelagic forum 2019.Rmd
#
# 06/07/2018 first coding
# 14/12/2018 updated for the new database structure
# 01/06/2019 updated for herring advice round
# 29/09/2019 adapted from stock summary code
# 03/10/2019 converted to powerpoint presentation
# -----------------------------------------------------------------------------------------------

if (packageVersion("rmarkdown") <= "1.9") {
  install.packages("rmarkdown")  # update rmarkdown from CRAN
}

rmarkdown::pandoc_version()

require("knitr")
knitr::opts_chunk$set(echo = FALSE,	message = FALSE,	warning = FALSE,	comment = "",	crop = TRUE )
knitr::opts_chunk$set(fig.width=10) 
knit_hooks$set(crop = hook_pdfcrop)

library(tidyverse)
library(cowplot)
library(directlabels)  # for printing labels at end of geom lines
library(scales)
library(stringr)
library(readxl)
library(gridExtra)                   # for plotting table as image (bio data)
library(RColorBrewer)

# Load utils code
source("../../mptools/r/my_utils.r")
source("iPlot_utils.r")

# Set dropbox folder
dropboxdir <- paste(get_dropbox(), "/iAdvice", sep="")

# load the data
iSpecies <-
  get(load(file=paste(dropboxdir,"/rdata/iSpecies.RData", sep=""))) %>% 
  ungroup() %>% 
  dplyr::select(speciesfaocode, trophicguild, fisheriesguild, sizeguild)

iAssess <-
  get(load(file=paste(dropboxdir,"/rdata/iAssess.RData", sep=""))) %>% 
  left_join(iSpecies, by="speciesfaocode")

load(file=paste(dropboxdir,"/rdata/iAdvice.RData", sep=""))
load(file=paste(dropboxdir,"/rdata/iRename.RData", sep=""))
load(file=paste(dropboxdir,"/rdata/iStockkey.RData", sep=""))

# add intermediate and forecast data
forecasts <-
  read_excel(path=paste(dropboxdir,"/excel/Pelagic forecasts 2019.xlsx", sep=""), col_names = TRUE) %>%
  left_join(iStockkey, by="stockkeylabelold") %>%
  mutate(stockkeylabel = stockkeylabelnew) %>%
  left_join(iRename,   by=c("stockkey", "stockkeylabel")) %>%
  left_join(iSpecies,  by="speciesfaocode") %>%
  mutate(catcheslandingsunits = "tonnes")

# t <-
#   icesSAGfull %>%
#   filter(year < assessmentyear) %>% 
#   bind_rows(forecasts)

fay <- 2011
fy  <- 1990
lay <- 2019
ly  <- 2019

```

# In the morning

## Catches by fisheries guild

```{r, echo=FALSE, fig.asp=0.5, fig.align="center", message=FALSE, warning=FALSE}

# catches by fisheriesguild

d <-
  iAssess %>% 
  ungroup() %>% 
  # filter(assessmentyear      == lay | (assessmentyear == 2017 & stockkeylabelold == "cap-icel") ) %>% 
  filter(assessmentyear      == lay  ) %>% 
  # filter(catcheslandingsunits == "tonnes") %>% 
  filter(year >= fy, year < ly) %>% 
  filter(purpose %in% c("advice")) %>%

  mutate(catches = ifelse(is.na(catches) | catches == 0, landings, catches))  %>% 
  mutate(fisheriesguild = ifelse(fisheriesguild %in% c("pelagic","demersal"), fisheriesguild, "other"))  %>% 
  mutate(fisheriesguild = factor(fisheriesguild, levels=c("pelagic","demersal","other"))) %>%

  group_by(fisheriesguild, speciesfaocode, year) %>% 
  summarize(catches  = sum(catches, na.rm=TRUE),
            landings = sum(landings, na.rm=TRUE)) %>%

  group_by(fisheriesguild, year) %>% 
  summarize(catches  = sum(catches, na.rm=TRUE),
            landings = sum(landings, na.rm=TRUE)) %>% 
  group_by(year) %>% 
  mutate(perc = round(100 * catches/ sum(catches), digits=2)) %>% 
  arrange(catches) %>% 
  mutate(fisheriesguild = factor(fisheriesguild, fisheriesguild))

# d %>% filter(year == 2018) %>% arrange(-catches) %>% View()

d2 <-
  d %>% 
  group_by(fisheriesguild) %>% 
  summarize(perc = mean(perc),
            catches=mean(catches),
            year = mean(year)) %>% 
  ungroup() %>% 
  mutate(cumcatches=cumsum(catches))

# plot stacked catches: pelagic, demersal and other
p1 <-
  d %>%

  ggplot(aes(year,catches)) +
  theme_publication() +
  theme(legend.position="none") +
  geom_bar(aes(fill=(fisheriesguild)), stat="identity", position="stack", colour="black") +
  # geom_text(data=subset(d, year==ly-1), aes(label=round(perc,0), group=fisheriesguild), 
  #           position = position_stack(vjust = 0.5), colour="white") +
  geom_text(data=subset(d, year==ly-1), 
            aes(label=fisheriesguild, group=fisheriesguild, x=ly), 
            position = position_stack(vjust = 0.5), colour="black", hjust=0) +
  # geom_dl(aes(label= perc),
  #         position = position_stack(vjust = 0.5),
  #         method   = list(dl.trans(x=x+0.9), dl.combine("last.points"),
  #                         fontface = 'bold',cex=.9, hjust = 0)) +
  scale_fill_brewer(palette = "Paired") +
  labs(x="",y="tonnes",title="catches") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(limits=c(fy,ly+3))

# d %>% filter(year == 2018) %>% View()

# plot of filled relative catches (how to add the direct labels?)
p2 <-
  d %>% 
  ggplot(aes(year,perc)) +
  theme_publication() +
  theme(legend.position="none") +
  theme(plot.margin = unit(c(0,0,0,0), "cm")) +
  theme(plot.background = element_blank()) +
  theme(axis.text = element_blank()) +
  theme(axis.ticks = element_blank()) +
  geom_bar(aes(fill=fisheriesguild), stat="identity", position="stack", colour="black") +
  scale_fill_brewer(palette = "Paired") +
  labs(x="",y="",title="") +
  # scale_y_continuous(labels = comma) +
  scale_x_continuous(limits=c(fy,ly))

vp <- grid::viewport(width = 0.3, height = 0.3, 
                     x = 0.15, y = 0.12, 
                     just = c("left", "bottom"))

 print(p1)
     theme_set(theme_bw(base_size = 8))
     # theme_white()
     print(p2, vp = vp)
     theme_set(theme_bw())

  # cowplot::plot_grid(p1, p2,
  #                ncol=2, scale=c(1.0, 0.25), align="hv",
  #                rel_widths = c(1.0, 1.0), rel_heights = c(1.0, 1.0))

```

## More information

Please refer to the [ICES website](http://www.ices.dk/community/advisory-process/Pages/Latest-advice.aspx) for an overview of all advice documents.