### 2022-12-17 Rmd/ICES assessments averaged.Rmd

---
output: 
  word_document:
    reference_docx: ../report_template_v1.5.dotx
---

```{r, message=FALSE, warning=FALSE, include=FALSE}

# -----------------------------------------------------------------------------------------------
# ICES assessments averaged.Rmd
#
# 06/07/2018 first coding
# 14/12/2018 updated for the new database structure
# 01/06/2019 updated for herring advice round
# -----------------------------------------------------------------------------------------------

require("knitr")
knitr::opts_chunk$set(echo = FALSE,	message = FALSE,	warning = FALSE,	comment = "",	crop = TRUE )
knitr::opts_chunk$set(fig.width=10) 
knit_hooks$set(crop = hook_pdfcrop)

# Libraries
library(rmarkdown) # creating markdown documents
library(tidyverse) # combined package of dplyr, tidyr, ggplot, readr, purrr and tibble
library(reshape2)  # reshaping data; e.g. dcast
library(ggthemes)  # for themes
library(pander)    # for print tables
library(readxl)    # read excel files
library(cowplot)   # multiplots
library(directlabels)  # label graphs
library(RColorBrewer) # colours
library(lubridate)

# Load utils code
source("../../prf/r/my utils.r")

dropboxdir <- paste(get_dropbox(), "/iAdvice", sep="")

# Load dataset
load(file=paste(dropboxdir, "/rdata/iAssess.RData",sep=""))
load(file=paste(dropboxdir, "/rdata/iAdvice.RData",sep=""))

```

**Summary of ICES 2019 assessments and advice**

Martin Pastoors, `r format(Sys.time(), '%d/%m/%Y')`

```{r, echo=FALSE, fig.asp=1.2, fig.align="center", message=FALSE, warning=FALSE}

iAssess %>% 
  filter(assessmentyear==2022) %>% 
  filter(year >= 1980, year < assessmentyear) %>% 
  filter(stocksizeunits =="tonnes") %>%

  group_by(stockkey) %>% 
  filter(n() >= 41) %>%

  # filter(grepl("her", stockkeylabelnew)) %>% 
  # distinct(stockkeylabelnew) %>% write.csv(file="C:/TEMP/stocks.csv", row.names = FALSE)

  group_by(year) %>% 
  summarise(
    stocksize       = sum(stocksize, na.rm=TRUE), 
    fishingpressure = mean(fishingpressure, na.rm=TRUE), 
    catches           = sum(catches, na.rm=TRUE)
  ) %>% 
  pivot_longer(names_to = "variable", values_to = "data", stocksize:catches) %>%

  # filter(year >= 2020) %>% View()

  ggplot(aes(x=year, y=data)) +
  theme_publication() +
  geom_line(aes()) +
  expand_limits(y=0) +
  facet_wrap(~variable, scale="free_y", ncol=1)

```

\newpage

```{r, echo=FALSE, fig.asp=1.2, fig.align="center", message=FALSE, warning=FALSE}

iAssess %>% 
  filter(assessmentyear==2022) %>% 
  filter(year >= 1980, year < assessmentyear) %>% 
  filter(stocksizeunits =="tonnes") %>%

  group_by(stockkey) %>% 
  filter(n() >= 41) %>%

  mutate(
    speciestype = case_when(
      grepl("cod|had|whg|sai|pol", stockkeylabelnew) ~ "roundfish",
      grepl("her|mac|whb|hom|anc|pil", stockkeylabelnew) ~ "pelagic",
      grepl("ple|sol|meg|tur|bll", stockkeylabelnew) ~ "flatfish",
      TRUE ~ as.character(NA)
    )
  ) %>% 
  drop_na(speciestype) %>%

  group_by(speciestype, year) %>% 
  summarise(
    stocksize       = sum(stocksize, na.rm=TRUE), 
    fishingpressure = mean(fishingpressure, na.rm=TRUE), 
    catches           = sum(catches, na.rm=TRUE)
  ) %>% 
  pivot_longer(names_to = "variable", values_to = "data", stocksize:catches) %>%

  # filter(year >= 2020) %>% View()

  ggplot(aes(x=year, y=data)) +
  theme_publication() +
  geom_line(aes()) +
  expand_limits(y=0) +
  facet_wrap(~speciestype+variable, scale="free_y", ncol=3)

```

```{r, echo=FALSE, fig.asp=1.2, fig.align="center", message=FALSE, warning=FALSE}

t <-
  iAssess %>% 
  filter(assessmentyear==2022) %>% 
  filter(year >= 1980, year < assessmentyear) %>% 
  filter(stocksizeunits =="tonnes") %>% 
  filter(purpose=="advice") %>%

  group_by(stockkey) %>% 
  filter(n() >= 41) %>%

  mutate(
    speciestype = case_when(
      grepl("cod|had|whg|sai|pol", stockkeylabelnew) ~ "roundfish",
      grepl("her|mac|whb|hom|anc|pil", stockkeylabelnew) ~ "pelagic",
      grepl("ple|sol|meg|tur|bll", stockkeylabelnew) ~ "flatfish",
      TRUE ~ as.character(NA)
    )
  ) %>% 
  drop_na(speciestype)

t %>% 
  dplyr::select(speciestype, stockkeylabelnew, year, stocksize, catches) %>% 
  pivot_longer(names_to = "variable", values_to = "data", stocksize:catches) %>%

  ggplot(aes(x=year, y=data, group=stockkeylabelnew)) +
  theme_publication() +
  geom_bar(aes(fill=stockkeylabelnew), stat="identity") +
  facet_wrap(~speciestype+variable, scales="free_y", ncol=2)

```