### 2022-12-17 Rmd/Metadata overview.Rmd

---
output: 
  word_document:
    reference_docx: ../report_template_v1.5.dotx
---

```{r, message=FALSE, warning=FALSE, include=FALSE}

# -----------------------------------------------------------------------------------------------
# Metadata overview.Rmd
#
# 14/11/2022 first coding
# -----------------------------------------------------------------------------------------------

require("knitr")
knitr::opts_chunk$set(echo = FALSE,	message = FALSE,	warning = FALSE,	comment = "",	crop = TRUE )
knitr::opts_chunk$set(fig.width=10) 
knit_hooks$set(crop = hook_pdfcrop)

# Libraries
library(tidyverse) # for piping and easy coding
library(reshape2)  # reshaping data; e.g. dcast
library(ggthemes)  # for themes
# library(ggExtra)   # for specific type of plots e.g. ggMarginal
library(pander)    # for print tables
library(readxl)    # read excel files
library(cowplot)   # multiplots
library(RColorBrewer) # colours
library(lubridate)
library(scales)

# captioner library
library(captioner) # deal with captions
tab_nums <- captioner::captioner(prefix = "Table", levels=2, type=c("n","n"), infix=".")
fig_nums <- captioner::captioner(prefix = "Figure", levels=2, type=c("n","n"), infix=".")

# Load utils code
source("../../prf/r/my utils.r")

dropboxdir <- paste(get_dropbox(), "/iAdvice", sep="")

# Load dataset
load(file=paste(dropboxdir, "/rdata/iAssess.RData",sep=""))
load(file=paste(dropboxdir, "/rdata/iAdvice.RData",sep=""))

# Create subset of series (stocks) to be used for analysis
d1 <-
  iAdvice %>%
  filter(adviceonstock == TRUE) %>% 
  group_by(speciescommonname, stockkey, stockkeylabelnew, stockkeylabelold) %>% 
  summarise(
    nyears          = n_distinct(tacyear),
    nadvicelandings = sum(!is.na(advisedlandingsmax)),
    nadvicecatch    = sum(!is.na(advisedcatchmax)),
    nblim           = sum(!is.na(blim)),
    nfset           = sum(!is.na(fset)),
    nssbset         = sum(!is.na(ssbset)),
  ) %>%

  filter(nssbset > 0)

  # filter stocks
d2 <-
  iAdvice %>%
  filter(stockkey %in% d1$stockkey) %>% 
  filter(adviceonstock == TRUE) %>% 
  group_by(speciescommonname, stockkey, stockkeylabelnew, stockkeylabelold, tacyear) %>% 
  summarise(
    nblim           = sum(!is.na(blim)),
    nfmsy           = sum(!is.na(fmsy)),
    nadvicelandings = sum(!is.na(advisedlandingsmax)),
    nadvicecatch    = sum(!is.na(advisedcatchmax)),
    nfset           = sum(!is.na(fset)),
    nssbset         = sum(!is.na(ssbset)),
    ntal            = sum(!is.na(tal)),
    ntac            = sum(!is.na(tac)),
    nlandings       = sum(!is.na(landings)),
    ncatch          = sum(!is.na(catches))
  )

d2 %>% 
  tidyr::pivot_longer(names_to = "variable", values_to = "data", nblim:ncatch) %>% 
  filter(data > 0) %>% 
  mutate(
    variable = factor(variable,
                      levels =c("nblim","nfmsy","nadvicelandings","nadvicecatch", "nfset", "nssbset",
                                "ntal","ntac","nlandings","ncatch"))
  ) %>%

  ggplot(aes(x=tacyear, y=variable)) +
  theme_publication() +
  geom_point() +
  facet_wrap(~stockkeylabelold)

```

<!-- Plot assessment and reconstructions      ------------------------------------------------------ -->

```{r, echo=FALSE, fig.asp=0.7, fig.align="center", message=FALSE, warning=FALSE}

# Create subset of series (stocks) to be used for analysis
a <-
  iAssess %>%
  filter(stockkey %in% d1$stockkey) %>% 
  group_by(stockkey) %>% 
  filter(assessmentyear == max(assessmentyear, na.rm=TRUE)) %>% 
  mutate(catch = ifelse(is.na(catches), landings, catches)) %>% 
  dplyr::select(
    stockkey, stockkeylabelnew, stockkeylabelold, purpose, assessmentyear, year, 
    lowstocksize, stocksize, highstocksize,
    lowfishingpressure, fishingpressure, highfishingpressure,
    catch
  ) %>% 
  tidyr::pivot_longer(names_to = "variable", values_to = "data", lowstocksize:catch) %>% 
  mutate(
    metric = case_when(
      grepl("low", variable)  ~ "low",
      grepl("high", variable) ~ "high", 
      TRUE                    ~ "est"
    )
  ) %>% 
  mutate(variable = gsub("low|high", "", variable)) %>% 
  pivot_wider(names_from = metric, values_from = data)

d <-
  iAdvice %>%
  filter(stockkey %in% d1$stockkey) %>% 
  filter(adviceonstock == TRUE) %>% 
  filter(!is.na(fset) | !is.na(ssbset)) %>% 
  dplyr::select(speciescommonname, stockkey, stockkeylabelnew, stockkeylabelold, year=tacyear, 
                fishingpressure=fset, stocksize=ssbset, catch=catches) %>% 
  tidyr::pivot_longer(names_to = "variable", values_to = "est", fishingpressure:catch)

for(myvar in c("stocksize", "fishingpressure", "catch")) {
  p <-
    ggplot() +
    theme_publication() +
    geom_point(data=filter(a, variable == myvar, year >= 2000), 
               aes(x=year, y=est)) +
    geom_line(data=filter(a, variable == myvar, year >= 2000), 
               aes(x=year, y=est)) +

    geom_point(data=filter(d, variable == myvar, year >= 2000), 
               aes(x=year, y=est),
               colour="red") +
    expand_limits(y=0) +
    labs(title=myvar) +
    facet_wrap(~stockkeylabelold, scales="free_y")

  print(p)
}

```

```{r, echo=FALSE, fig.asp=0.7, fig.align="center", message=FALSE, warning=FALSE}

c <-
  a %>% 
  mutate(source="assessed") %>% 
  dplyr::select(-purpose, -assessmentyear) %>% 
  bind_rows(
    d %>% 
      mutate(source="predicted")
  ) %>% 
  dplyr::select(-low, -high, -speciescommonname) %>% 
  tidyr::pivot_wider(names_from = source, values_from = est) %>% 
  drop_na(predicted)

myvar <- "stocksize"

for(myvar in c("stocksize", "fishingpressure", "catch")) {
  p <-
    c %>% 
    filter(variable == myvar) %>%

    ggplot(aes(x=predicted, y=assessed)) +
    theme_publication() +
    geom_point() +
    geom_abline(slope=1, intercept=0) +
    expand_limits(x=0, y=0) +
    labs(title=myvar) +
    facet_wrap(~stockkeylabelold, scales="free")

  print(p)
}

```