### 2023-07-19 Rmd/Paper 03 short term forecast analysis.Rmd
Several updates

---
output: 
  word_document:
    reference_docx: ../report_template_v1.5.dotx
---

```{r, message=FALSE, warning=FALSE, include=FALSE}

# -----------------------------------------------------------------------------------------------
# Paper 03 short term forecast analysis.Rmd
#
# 17/12/2022 first coding
# -----------------------------------------------------------------------------------------------

require("knitr")
knitr::opts_chunk$set(echo = FALSE,	message = FALSE,	warning = FALSE,	comment = "",	crop = TRUE )
knitr::opts_chunk$set(fig.width=10) 
knit_hooks$set(crop = hook_pdfcrop)

# Libraries
library(rmarkdown) # creating markdown documents
library(tidyverse) # combined package of dplyr, tidyr, ggplot, readr, purrr and tibble
library(reshape2)  # reshaping data; e.g. dcast
library(ggthemes)  # for themes
library(pander)    # for print tables
library(readxl)    # read excel files
library(cowplot)   # multiplots
library(directlabels)  # label graphs
library(RColorBrewer) # colours
library(lubridate)
library(ggrepel)
library(patchwork)

# Load utils code
source("../../prf/r/my utils.r")
source("../r/iAssess functions.r")

dropboxdir <- paste(get_dropbox(), "/iAdvice", sep="")

mystocks <- c("her-47d3", "whb-comb", "mac-nea", "hom-west", 
              "ple-nsea", "cod-nsea", "sol-nsea","had-nsea", "hke-nrtn")

demstocks <- c("ple-nsea", "cod-nsea", "sol-nsea","had-nsea", "hke-nrtn")
pelstocks <- c("her-47d3", "whb-comb", "mac-nea", "hom-west")

# Load dataset
iAssess <-
  loadRData(file=paste(dropboxdir, "/rdata/iAssess.RData",sep="")) %>% 
  mutate(stockkeylabelold = ifelse(stockkeylabelold=="ple-nsea2", 
                                   "ple-nsea", 
                                   stockkeylabelold)) %>% 
  mutate(stockkeylabelold = ifelse(stockkeylabelold %in% c("had-nsea","had-34","had-346a"),
                                   "had-nsea",
                                   stockkeylabelold)) %>% 
  mutate(stockkeylabelold = ifelse(stockkeylabelold %in% c("cod-nsea","cod-347d"),
                                   "cod-nsea",
                                   stockkeylabelold)) %>% 
  filter(stockkeylabelold %in% mystocks)

# iAdvice %>% ungroup() %>% arrange(stockkeylabelold) %>% distinct(stockkeylabelold) %>% View()

iAdvice <-
  loadRData(file=paste(dropboxdir, "/rdata/iAdvice.RData",sep="")) %>% 
  mutate(stockkeylabelold = ifelse(stockkeylabelold=="ple-nsea2", 
                                   "ple-nsea", 
                                   stockkeylabelold)) %>% 
  mutate(stockkeylabelold = ifelse(stockkeylabelold %in% c("had-nsea","had-34","had-346a"),
                                   "had-nsea",
                                   stockkeylabelold)) %>% 
  mutate(stockkeylabelold = ifelse(stockkeylabelold %in% c("cod-nsea","cod-347d"),
                                   "cod-nsea",
                                   stockkeylabelold)) %>% 
  filter(stockkeylabelold %in% mystocks)

# iAdvice %>% ungroup() %>% distinct(stockkeylabelold) %>% filter(grepl("had-",stockkeylabelold)) %>% View()

fay <- 1990
fy  <- 1990
lay <- 2022
ly  <- 2021

assessmentyears         <- c(2015, 2020, 2022)

myColors                <- RColorBrewer::brewer.pal(12, "Paired")[c(2:4,6,8,10)]
names(myColors)         <- c("1990", "2000", "2010", "2020", "2022")

# iAdvice %>% distinct(stockkeylabelold, assessmentyear) %>% filter(stockkeylabelold=="hke-nrtn") %>% View()
# iAssess %>% distinct(stockkeylabelold, assessmentyear) %>% filter(stockkeylabelold=="hke-nrtn") %>% View()

# iAssess %>% filter(stockkeylabelold=="hke-nrtn", assessmentyear==2010) %>% View()

```

**Paper 03 Short term forecast analysis**

Martin Pastoors, `r format(Sys.time(), '%d/%m/%Y')`

<!--Section 1: introduction ------------------------------------------------------------------- -->

# Introduction

<!--Section 2: material and methods ----------------------------------------------------------- -->

# Material and methods

<!--Section 3: advice and assessment overviews ------------------------------------------------ -->

# Assessment and advice overviews

```{r, echo=FALSE, fig.asp=1.4, fig.align="center", message=FALSE, warning=FALSE}

iAdvice %>% 
  group_by(stockkeylabelold, tacyear) %>% 
  summarize(
    nF = sum(!is.na(fset)),
    nSSB = sum(!is.na(ssbset))
  ) %>% 
  tidyr::pivot_longer(names_to = "variable", values_to = "data", nF:nSSB) %>% 
  # mutate(data = ifelse(data==0, NA, data)) %>% 
  filter(data > 0) %>% 
  ggplot(aes(x=tacyear, y=variable)) +
  theme_publication() +
  geom_point(aes(colour=variable)) +
  facet_wrap(~stockkeylabelold)

```

**Demersal stocks**

```{r, echo=FALSE, fig.asp=1.4, fig.align="center", message=FALSE, warning=FALSE}

## Demersal

# advice
d1 <-
  iAdvice %>%
  filter(stockkeylabelold %in% demstocks) %>%
  filter(adviceonstock) %>%
  filter(tacyear >= fy, tacyear <= ly) %>%
  dplyr::select(stockkeylabelold, year=tacyear, 
                f=fset, ssb=ssbset) %>%
  mutate(decade = as.character(10 * floor(year/10))) %>% 
  mutate(yr = stringr::str_pad(ifelse(year >=2000, year-2000, year-1900), 
                               width=2, pad="0")) %>% 
  tidyr::pivot_longer(names_to = "variable", values_to = "data", ssb:f) %>% 
  mutate(variable = factor(variable, levels=c("ssb","f","rec")))

# assess
d2 <-
  iAssess %>%
  filter(stockkeylabelold %in% demstocks) %>%
  filter(purpose=="advice") %>% 
  # filter(assessmentyear == max(assessmentyear)) %>%
  filter(year >= fy, year <= ly) %>%
  dplyr::select(stockkeylabelold, assessmentyear, year, 
                ssb=stocksize, f=fishingpressure, rec=recruitment) %>%
  mutate(decade = as.character(10 * floor(assessmentyear/10))) %>% 
  tidyr::pivot_longer(names_to = "variable", values_to = "data", ssb:rec) %>% 
  mutate(variable = factor(variable, levels=c("ssb","f","rec")))

i <- 1
for (i in 1:length(demstocks)) {

  assign(paste0("p",i),
         d2 %>% 
           filter(stockkeylabelold == demstocks[[i]]) %>%

           ggplot(aes(x=year, y=data, group=assessmentyear)) +
           theme_publication() +
           geom_line(aes(colour=decade)) +
           geom_point(data=filter(d1, stockkeylabelold == demstocks[[i]]), 
                      aes(x=year, y=data, colour=decade),
                      inherit.aes=FALSE) +
           ggrepel::geom_text_repel(
                      data=filter(d1, stockkeylabelold == demstocks[[i]]),
                      aes(x=year, y=data, colour=decade, label=yr),
                      inherit.aes=FALSE) +
           expand_limits(y=0) +
           scale_colour_manual(values=myColors) +
           labs(x="",y="", title=demstocks[[i]]) +
           facet_wrap(~variable, scales="free_y")
         )
}

p1 + p2 + p3 + p4 + p5 + plot_layout(ncol = 1, guides="collect") &
  theme(legend.position='bottom')

```

\newpage

**Pelagic stocks**

```{r, echo=FALSE, fig.asp=1.2, fig.align="center", message=FALSE, warning=FALSE}

## Pelagic

# advice
d1 <-
  iAdvice %>%
  filter(stockkeylabelold %in% pelstocks) %>%
  filter(adviceonstock) %>%
  filter(tacyear >= fy, tacyear <= ly) %>%
  dplyr::select(stockkeylabelold, year=tacyear, 
                f=fset, ssb=ssbset) %>%
  mutate(decade = as.character(10 * floor(year/10))) %>% 
  mutate(yr = stringr::str_pad(ifelse(year >=2000, year-2000, year-1900), 
                               width=2, pad="0")) %>% 
  tidyr::pivot_longer(names_to = "variable", values_to = "data", ssb:f) %>% 
  mutate(variable = factor(variable, levels=c("ssb","f","rec")))

# assess
d2 <-
  iAssess %>%
  filter(stockkeylabelold %in% pelstocks) %>%
  filter(purpose=="advice") %>% 
  # filter(assessmentyear == max(assessmentyear)) %>%
  filter(year >= fy, year <= ly) %>%
  dplyr::select(stockkeylabelold, assessmentyear, year, 
                ssb=stocksize, f=fishingpressure, rec=recruitment) %>%
  mutate(decade = as.character(10 * floor(assessmentyear/10))) %>% 
  tidyr::pivot_longer(names_to = "variable", values_to = "data", ssb:rec) %>% 
  mutate(variable = factor(variable, levels=c("ssb","f","rec")))

i <- 1
for (i in 1:length(pelstocks)) {

  assign(paste0("p",i),
         d2 %>% 
           filter(stockkeylabelold == pelstocks[[i]]) %>%

           ggplot(aes(x=year, y=data, group=assessmentyear)) +
           theme_publication() +
           geom_line(aes(colour=decade)) +
           geom_point(data=filter(d1, stockkeylabelold == pelstocks[[i]]), 
                      aes(x=year, y=data, colour=decade),
                      inherit.aes=FALSE) +
           ggrepel::geom_text_repel(
                      data=filter(d1, stockkeylabelold == pelstocks[[i]]),
                      aes(x=year, y=data, colour=decade, label=yr),
                      inherit.aes=FALSE) +
           scale_colour_manual(values=myColors) +
           expand_limits(y=0) +
           labs(x="",y="", title=pelstocks[[i]]) +
           facet_wrap(~variable, scales="free_y")
         )
}

p1 + p2 + p3 + p4 + plot_layout(ncol = 1, guides="collect") &
  theme(legend.position='bottom')

```

\newpage

**Demersal F**

```{r, echo=FALSE, fig.asp=1.4, fig.align="center", message=FALSE, warning=FALSE}

# plot of F against Fset and SSB against SSBset for multiple assessments

mytype    <- "Demersal"
myvar     <- "f"

d1 <-
  iAdvice %>%
  filter(stockkeylabelold %in% demstocks) %>%
  filter(adviceonstock) %>%
  filter(tacyear >= fay, tacyear <= lay) %>%
  dplyr::select(stockkeylabelold, year=tacyear, f=fset, ssb=ssbset) %>%
  tidyr::pivot_longer(names_to = "variable", values_to = "data", ssb:f)

d2 <-
  iAssess %>%
  filter(stockkeylabelold %in% demstocks) %>%
  filter(purpose=="advice") %>% 
  filter(assessmentyear %in% assessmentyears) %>% 
  filter(year >= fay, year <= lay) %>%
  mutate(source=paste0("assess", assessmentyear)) %>% 
  dplyr::select(stockkeylabelold, year, ssb=stocksize, f=fishingpressure, source) %>%
  tidyr::pivot_longer(names_to = "variable", values_to = "data", ssb:f)

# iAssess %>% filter(stockkeylabelold=="hke-nrtn", assessmentyear==2010) %>% View()

t <-
  d2 %>% 
  left_join(d1, by=c("stockkeylabelold", "year", "variable")) %>% 
  rename(
    assess=data.x,
    forecast=data.y
  ) %>% 
  drop_na(forecast) %>% 
  mutate(decade = as.character(10 * floor(year / 10))) %>% 
  mutate(yr = stringr::str_pad(ifelse(year >=2000, year-2000, year-1900), width=2, pad="0"))

# maxssb<- t %>% filter(variable=="ssb") %>% ungroup() %>% 
#   summarise(maxssb=max(assess, forecast, na.rm=TRUE)) %>% 
#   as.numeric()

i <- 1
for (i in 1:length(demstocks)) {

  max  <- t %>% 
    filter(variable==myvar) %>% 
    filter(stockkeylabelold == demstocks[[i]]) %>% 
    ungroup() %>%
    summarise(max=max(assess, forecast, na.rm=TRUE)) %>%
    as.numeric()

  assign(paste0("p",i),
         t %>% 
           filter(variable == myvar) %>% 
           filter(stockkeylabelold == demstocks[[i]]) %>%

           ggplot(aes(x=assess, y=forecast)) +
           theme_publication() +
           geom_point(aes(colour=decade)) +
           ggrepel::geom_text_repel(aes(colour=decade, label=yr)) +
           scale_colour_manual(values=myColors) +
           geom_abline(slope=1, intercept=0, linetype="dashed", linewidth=0.5) +
           scale_x_continuous(limits=c(0,max)) +
           scale_y_continuous(limits=c(0,max)) +
           facet_grid(stockkeylabelold~source)
         )
}

p1 + p2 + p3 + p4 + p5 + plot_layout(ncol = 1, guides="collect") &
  theme(legend.position='bottom')

```

\newpage

**Demersal SSB**

```{r, echo=FALSE, fig.asp=1.4, fig.align="center", message=FALSE, warning=FALSE}

# plot of F against Fset and SSB against SSBset for multiple assessments

mytype    <- "Demersal"
myvar     <- "ssb"

d1 <-
  iAdvice %>%
  filter(stockkeylabelold %in% demstocks) %>%
  filter(adviceonstock) %>%
  filter(tacyear >= fay, tacyear <= lay) %>%
  dplyr::select(stockkeylabelold, year=tacyear, f=fset, ssb=ssbset) %>%
  tidyr::pivot_longer(names_to = "variable", values_to = "data", ssb:f)

d2 <-
  iAssess %>%
  filter(stockkeylabelold %in% demstocks) %>%
  filter(purpose=="advice") %>% 
  filter(assessmentyear %in% assessmentyears) %>% 
  filter(year >= fay, year <= lay) %>%
  mutate(source=paste0("assess", assessmentyear)) %>% 
  dplyr::select(stockkeylabelold, year, ssb=stocksize, f=fishingpressure, source) %>%
  tidyr::pivot_longer(names_to = "variable", values_to = "data", ssb:f)

# d2 %>% filter(stockkeylabelold=="hke-nrtn", source=="assess1990") %>% View()

t <-
  d2 %>% 
  left_join(d1, by=c("stockkeylabelold", "year", "variable")) %>% 
  rename(
    assess=data.x,
    forecast=data.y
  ) %>% 
  drop_na(forecast) %>% 
  mutate(decade = as.character(10 * floor(year / 10))) %>% 
  mutate(yr = stringr::str_pad(ifelse(year >=2000, year-2000, year-1900), width=2, pad="0"))

# maxssb<- t %>% filter(variable=="ssb") %>% ungroup() %>% 
#   summarise(maxssb=max(assess, forecast, na.rm=TRUE)) %>% 
#   as.numeric()

i <- 1
for (i in 1:length(demstocks)) {

  max  <- t %>% 
    filter(variable==myvar) %>% 
    filter(stockkeylabelold == demstocks[[i]]) %>% 
    ungroup() %>%
    summarise(max=max(assess, forecast, na.rm=TRUE)) %>%
    as.numeric()

  assign(paste0("p",i),
         t %>% 
           filter(variable == myvar) %>% 
           filter(stockkeylabelold == demstocks[[i]]) %>%

           ggplot(aes(x=assess, y=forecast)) +
           theme_publication() +
           geom_point(aes(colour=decade)) +
           ggrepel::geom_text_repel(aes(colour=decade, label=yr)) +
           scale_colour_manual(values=myColors) +
           geom_abline(slope=1, intercept=0, linetype="dashed", linewidth=0.5) +
           scale_x_continuous(limits=c(0,max)) +
           scale_y_continuous(limits=c(0,max)) +
           facet_grid(stockkeylabelold~source)
         )
}

p1 + p2 + p3 + p4 + p5 + plot_layout(ncol = 1, guides="collect") &
  theme(legend.position='bottom')

```

\newpage

**Pelagic F**

```{r, echo=FALSE, fig.asp=1.2, fig.align="center", message=FALSE, warning=FALSE}

# plot of F against Fset and SSB against SSBset for multiple assessments

mytype    <- "Pelagic"
myvar     <- "f"

d1 <-
  iAdvice %>%
  filter(stockkeylabelold %in% pelstocks) %>%
  filter(adviceonstock) %>%
  filter(tacyear >= fay, tacyear <= lay) %>%
  dplyr::select(stockkeylabelold, year=tacyear, f=fset, ssb=ssbset) %>%
  tidyr::pivot_longer(names_to = "variable", values_to = "data", ssb:f)

d2 <-
  iAssess %>%
  filter(stockkeylabelold %in% pelstocks) %>%
  filter(purpose=="advice") %>% 
  filter(assessmentyear %in% assessmentyears) %>% 
  filter(year >= fay, year <= lay) %>%
  mutate(source=paste0("assess", assessmentyear)) %>% 
  dplyr::select(stockkeylabelold, year, ssb=stocksize, f=fishingpressure, source) %>%
  tidyr::pivot_longer(names_to = "variable", values_to = "data", ssb:f)

t <-
  d2 %>% 
  left_join(d1, by=c("stockkeylabelold", "year", "variable")) %>% 
  rename(
    assess=data.x,
    forecast=data.y
  ) %>% 
  drop_na(forecast) %>% 
  mutate(decade = as.character(10 * floor(year / 10))) %>% 
  mutate(yr = stringr::str_pad(ifelse(year >=2000, year-2000, year-1900), width=2, pad="0"))

# maxssb<- t %>% filter(variable=="ssb") %>% ungroup() %>% 
#   summarise(maxssb=max(assess, forecast, na.rm=TRUE)) %>% 
#   as.numeric()

i <- 1
for (i in 1:length(pelstocks)) {

  max  <- t %>% 
    filter(variable==myvar) %>% 
    filter(stockkeylabelold == pelstocks[[i]]) %>% 
    ungroup() %>%
    summarise(max=max(assess, forecast, na.rm=TRUE)) %>%
    as.numeric()

  assign(paste0("p",i),
         t %>% 
           filter(variable == myvar) %>% 
           filter(stockkeylabelold == pelstocks[[i]]) %>%

           ggplot(aes(x=assess, y=forecast)) +
           theme_publication() +
           geom_point(aes(colour=decade)) +
           ggrepel::geom_text_repel(aes(colour=decade, label=yr)) +
           scale_colour_manual(values=myColors) +
           geom_abline(slope=1, intercept=0, linetype="dashed", linewidth=0.5) +
           scale_x_continuous(limits=c(0,max)) +
           scale_y_continuous(limits=c(0,max)) +
           facet_grid(stockkeylabelold~source)
         )
}

p1 + p2 + p3 + p4 + plot_layout(ncol = 1, guides="collect") &
  theme(legend.position='bottom')

```

\newpage

**Pelagic SSB**

```{r, echo=FALSE, fig.asp=1.2, fig.align="center", message=FALSE, warning=FALSE}

# plot of F against Fset and SSB against SSBset for multiple assessments

mytype    <- "Pelagic"
myvar     <- "ssb"

d1 <-
  iAdvice %>%
  filter(stockkeylabelold %in% pelstocks) %>%
  filter(adviceonstock) %>%
  filter(tacyear >= fay, tacyear <= lay) %>%
  dplyr::select(stockkeylabelold, year=tacyear, f=fset, ssb=ssbset) %>%
  tidyr::pivot_longer(names_to = "variable", values_to = "data", ssb:f)

d2 <-
  iAssess %>%
  filter(stockkeylabelold %in% pelstocks) %>%
  filter(purpose=="advice") %>% 
  filter(assessmentyear %in% assessmentyears) %>% 
  filter(year >= fay, year <= lay) %>%
  mutate(source=paste0("assess", assessmentyear)) %>% 
  dplyr::select(stockkeylabelold, year, ssb=stocksize, f=fishingpressure, source) %>%
  tidyr::pivot_longer(names_to = "variable", values_to = "data", ssb:f)

t <-
  d2 %>% 
  left_join(d1, by=c("stockkeylabelold", "year", "variable")) %>% 
  rename(
    assess=data.x,
    forecast=data.y
  ) %>% 
  drop_na(forecast) %>% 
  mutate(decade = as.character(10 * floor(year / 10))) %>% 
  mutate(yr = stringr::str_pad(ifelse(year >=2000, year-2000, year-1900), width=2, pad="0"))

# maxssb<- t %>% filter(variable=="ssb") %>% ungroup() %>% 
#   summarise(maxssb=max(assess, forecast, na.rm=TRUE)) %>% 
#   as.numeric()

i <- 1
for (i in 1:length(pelstocks)) {

  max  <- t %>% 
    filter(variable==myvar) %>% 
    filter(stockkeylabelold == pelstocks[[i]]) %>% 
    ungroup() %>%
    summarise(max=max(assess, forecast, na.rm=TRUE)) %>%
    as.numeric()

  assign(paste0("p",i),
         t %>% 
           filter(variable == myvar) %>% 
           filter(stockkeylabelold == pelstocks[[i]]) %>%

           ggplot(aes(x=assess, y=forecast)) +
           theme_publication() +
           geom_point(aes(colour=decade)) +
           ggrepel::geom_text_repel(aes(colour=decade, label=yr)) +
           scale_colour_manual(values=myColors) +
           geom_abline(slope=1, intercept=0, linetype="dashed", linewidth=0.5) +
           scale_x_continuous(limits=c(0,max)) +
           scale_y_continuous(limits=c(0,max)) +
           facet_grid(stockkeylabelold~source)
         )
}

p1 + p2 + p3 + p4 + plot_layout(ncol = 1, guides="collect") &
  theme(legend.position='bottom')

```