### 2023-08-15 Rmd/Paper 03 short term forecast analysis v2.Rmd
UPdates STF paper

---
output: 
  word_document:
    reference_docx: ../report_template_v1.5.dotx
---

```{r, message=FALSE, warning=FALSE, include=FALSE}

# -----------------------------------------------------------------------------------------------
# Paper 03 short term forecast analysis.Rmd
#
# 17/12/2022 first coding
# August 2023 updated coding
# -----------------------------------------------------------------------------------------------

require("knitr")
knitr::opts_chunk$set(echo = FALSE,	message = FALSE,	warning = FALSE,	comment = "",	crop = TRUE )
knitr::opts_chunk$set(fig.width=10) 
knit_hooks$set(crop = hook_pdfcrop)

# Libraries
library(rmarkdown) # creating markdown documents
library(tidyverse) # combined package of dplyr, tidyr, ggplot, readr, purrr and tibble
library(reshape2)  # reshaping data; e.g. dcast
library(ggthemes)  # for themes
library(pander)    # for print tables
library(readxl)    # read excel files
library(cowplot)   # multiplots
library(directlabels)  # label graphs
library(RColorBrewer) # colours
library(lubridate)
library(ggrepel)
library(patchwork)

# Load utils code
source("../../prf/r/my utils.r")
source("../r/iAssess functions.r")

dropboxdir <- paste(get_dropbox(), "/iAdvice", sep="")

mystocks <- c("her-47d3", "whb-comb", "mac-nea", "hom-west", 
              "ple-nsea", "cod-nsea", "sol-nsea","had-nsea", "hke-nrtn")

demstocks <- c("ple-nsea", "cod-nsea", "sol-nsea","had-nsea", "hke-nrtn")
pelstocks <- c("her-47d3", "whb-comb", "mac-nea", "hom-west")

fay <- 1980
fy  <- 1980
lay <- 2023
ly  <- 2023

# Load dataset
iAssess <-
  loadRData(file=paste(dropboxdir, "/rdata/iAssess.RData",sep="")) %>% 
  mutate(stockkeylabelold = ifelse(stockkeylabelold=="ple-nsea2", 
                                   "ple-nsea", 
                                   stockkeylabelold)) %>% 
  mutate(stockkeylabelold = ifelse(stockkeylabelold %in% c("had-nsea","had-34","had-346a"),
                                   "had-nsea",
                                   stockkeylabelold)) %>% 
  mutate(stockkeylabelold = ifelse(stockkeylabelold %in% c("cod-nsea","cod-347d"),
                                   "cod-nsea",
                                   stockkeylabelold)) %>% 
  filter(purpose == "advice") %>% 
  filter(stockkeylabelold %in% mystocks) %>% 
  filter(year >= assessmentyear -9, year < assessmentyear) %>%

  mutate(decade = as.character(10 * floor(assessmentyear/10))) %>% 
  mutate(pentad = as.character(5 * floor(assessmentyear/5))) %>% 
  arrange(stockkeylabelold, assessmentyear, year) %>% 
  dplyr::select(stockkeylabelold, purpose, decade, pentad, assessmentyear, year, 
              ssb=stocksize, f=fishingpressure, rec=recruitment) %>%
  group_by(stockkeylabelold, pentad) %>% 
  filter(assessmentyear == max(assessmentyear, na.rm=TRUE)) %>% 
  ungroup()

# iAssess %>% filter(stockkeylabelold == "cod-nsea", pentad=="2010") %>% View()

# add intermediate year from iAdvice database
iAssess2 <-
  loadRData(file=paste(dropboxdir, "/rdata/iAdvice.RData",sep="")) %>% 
  mutate(stockkeylabelold = ifelse(stockkeylabelold=="ple-nsea2", 
                                   "ple-nsea", 
                                   stockkeylabelold)) %>% 
  mutate(stockkeylabelold = ifelse(stockkeylabelold %in% c("had-nsea","had-34","had-346a"),
                                   "had-nsea",
                                   stockkeylabelold)) %>% 
  mutate(stockkeylabelold = ifelse(stockkeylabelold %in% c("cod-nsea","cod-347d"),
                                   "cod-nsea",
                                   stockkeylabelold)) %>% 
  filter(adviceonstock == TRUE & purpose=="advice") %>% 
  filter(stockkeylabelold %in% mystocks) %>% 
  filter(tacyear >= fy, tacyear <= ly) %>%
  rename(ssb = ssbimy, f = fsqimy) %>% 
  mutate(year = assessmentyear + 1) %>% 
  mutate(pentad = as.character(5 * floor(year/5))) %>% 
  mutate(decade = as.character(10 * floor(year/10))) %>% 
  dplyr::select(stockkeylabelold, purpose, decade, pentad, assessmentyear, year, ssb, f)

# iAssess %>% ungroup() %>% arrange(stockkeylabelold) %>% distinct(stockkeylabelold, decade, assessmentyear) %>% View()
# iAssess %>% group_by(stockkeylabelold, decade, assessmentyear) %>% summarise(minyear = min(year), maxyear = max(year)) %>% View()

iAdvice <-
  loadRData(file=paste(dropboxdir, "/rdata/iAdvice.RData",sep="")) %>% 
  mutate(stockkeylabelold = ifelse(stockkeylabelold=="ple-nsea2", 
                                   "ple-nsea", 
                                   stockkeylabelold)) %>% 
  mutate(stockkeylabelold = ifelse(stockkeylabelold %in% c("had-nsea","had-34","had-346a"),
                                   "had-nsea",
                                   stockkeylabelold)) %>% 
  mutate(stockkeylabelold = ifelse(stockkeylabelold %in% c("cod-nsea","cod-347d"),
                                   "cod-nsea",
                                   stockkeylabelold)) %>% 
  filter(adviceonstock == TRUE & purpose == "advice") %>% 
  filter(stockkeylabelold %in% mystocks) %>% 
  filter(tacyear >= fy, tacyear <= ly) %>%

  rename(year = tacyear) %>%

  mutate(pentad = as.character(5 * floor(year/5))) %>% 
  mutate(decade = as.character(10 * floor(year/10))) %>% 
  arrange(stockkeylabelold, year) %>% 
  dplyr::select(stockkeylabelold, purpose, decade, pentad, assessmentyear, year, 
              ssb=ssbset, f=fset) %>%
  ungroup()

# iAdvice %>% ungroup() %>% distinct(stockkeylabelold) %>%  View()
# iAdvice %>% distinct(pentad) %>% View()
# iAdvice %>% filter(pentad==1975) %>% View()

# year strings
assessmentyears         <- iAdvice %>% distinct(pentad, year) %>% group_by(pentad) %>% filter(year == max(year)) %>% pull(year)
pentadyears             <- iAdvice %>% distinct(pentad) %>% pull(pentad)

# colours
myColors                <- RColorBrewer::brewer.pal(12, "Paired")[1:length(pentadyears)]
names(myColors)         <- pentadyears

# scales::show_col(RColorBrewer::brewer.pal(12, "Paired"))

# iAdvice %>% distinct(stockkeylabelold, assessmentyear) %>% filter(stockkeylabelold=="hke-nrtn") %>% View()

# iAssess %>% distinct(stockkeylabelold, assessmentyear) %>% filter(stockkeylabelold=="hke-nrtn") %>% View()
# iAssess %>% filter(stockkeylabelold=="hke-nrtn", assessmentyear==2010) %>% View()

```

**Paper 03 Short term forecast analysis**

Martin Pastoors, `r format(Sys.time(), '%d/%m/%Y')`

<!--Section 1: introduction ------------------------------------------------------------------- -->

# Introduction

<!--Section 2: material and methods ----------------------------------------------------------- -->

# Material and methods

<!--Section 3: advice and assessment overviews ------------------------------------------------ -->

# Assessment and advice overviews

Reconstructed F and SSB by stock, calculated from observed catches and implied F and SSB from short term forecasts in different years

```{r, echo=FALSE, fig.asp=1.4, fig.align="center", message=FALSE, warning=FALSE}

iAdvice %>% 
  group_by(stockkeylabelold, year) %>% 
  summarize(
    nSTF_F = sum(!is.na(f)),
    nSTF_SSB = sum(!is.na(ssb))
  ) %>% 
  tidyr::pivot_longer(names_to = "variable", values_to = "data", nSTF_F:nSTF_SSB) %>% 
  filter(data > 0) %>%

  bind_rows(
    iAssess %>% 
      ungroup() %>% 
      distinct(stockkeylabelold, assessmentyear) %>% 
      rename(year = assessmentyear) %>% 
      mutate(variable = "nAssess", data=1)
  ) %>%

  ggplot(aes(x=year, y=variable)) +
  theme_publication() +
  theme(legend.position = "none") +
  geom_point(aes(colour=variable)) +
  facet_wrap(~stockkeylabelold)

```

**Demersal stocks**

Historical retrospective assessments of demersal stocks (SSB)

```{r, echo=FALSE, fig.asp=1.4, fig.align="center", message=FALSE, warning=FALSE}

## Demersal
my_var <- "f"

# advice
d1 <-
  iAdvice %>%
  filter(stockkeylabelold %in% demstocks) %>%
  # filter(stockkeylabelold == "cod-nsea") %>% 
  mutate(yr = stringr::str_pad(ifelse(year >=2000, year-2000, year-1900), 
                               width=2, pad="0")) %>% 
  tidyr::pivot_longer(names_to = "variable", values_to = "data", ssb:f) %>% 
  mutate(variable = factor(variable, levels=c("ssb","f","rec"))) %>%

  filter(variable == my_var)

# assess
d2 <-
  iAssess %>%
  filter(stockkeylabelold %in% demstocks) %>%
  # filter(stockkeylabelold == "cod-nsea") %>%

  group_by(stockkeylabelold, pentad) %>% 
  filter(assessmentyear == max(assessmentyear)) %>%

  tidyr::pivot_longer(names_to = "variable", values_to = "data", ssb:rec) %>% 
  mutate(variable = factor(variable, levels=c("ssb","f","rec"))) %>%

  filter(variable == my_var)

# iAssess %>% filter(stockkeylabelold == "cod-nsea", pentad=="2010") %>% View()

# IMY assumptions for ssb and F
d3 <-
  iAssess2 %>%
  filter(stockkeylabelold %in% demstocks) %>%
  # filter(stockkeylabelold == "cod-nsea") %>%

  group_by(stockkeylabelold, pentad) %>% 
  filter(assessmentyear == max(assessmentyear)) %>%

  tidyr::pivot_longer(names_to = "variable", values_to = "data", ssb:f) %>% 
  mutate(variable = factor(variable, levels=c("ssb","f"))) %>%

  filter(variable == my_var)

# connecting last assessment point with IMY assumptions
d4 <-
  iAssess %>%
  filter(stockkeylabelold %in% demstocks) %>%
  # filter(stockkeylabelold == "cod-nsea") %>%

  group_by(stockkeylabelold, pentad) %>% 
  filter(assessmentyear == max(assessmentyear)) %>% 
  filter(year == max(year)) %>% 
  tidyr::pivot_longer(names_to = "variable", values_to = "data", ssb:f) %>% 
  mutate(variable = factor(variable, levels=c("ssb","f"))) %>%

  bind_rows(d3) %>%

  filter(variable == my_var)

int_breaks <- function(x, n = 5) {
  l <- pretty(x, n)
  l[abs(l %% 1) < .Machine$double.eps ^ 0.5] 
}

d2 %>% 
  ggplot(aes(x=year, y=data, group=assessmentyear)) +
  theme_publication() +
  theme(plot.margin = unit(c(0,0,0,0), "cm")) +
  theme(plot.title = element_text(size=12)) +
  theme(panel.spacing.x = unit(0.1, "mm"), panel.spacing.y = unit(0.1, "mm")) +
  theme(legend.position = "none") +

  geom_line(aes(colour=pentad), alpha=1) +
  ggrepel::geom_text_repel(
             data=d2 %>% group_by(stockkeylabelold, decade, pentad, assessmentyear) %>% filter(row_number()==1),
             aes(x=year, y=data, colour=pentad, label=assessmentyear),
             inherit.aes=FALSE) +
  geom_point(data=d3, 
             aes(x=year, y=data, colour=pentad),
             inherit.aes=FALSE, alpha=1, shape=1) +
  geom_line(data=d4, 
             aes(x=year, y=data, colour=pentad),
             inherit.aes=FALSE, alpha=1, linetype="dotted") +
  geom_point(data=d1,
             aes(x=year, y=data, colour=pentad),
             inherit.aes=FALSE, alpha=1) +
  ggrepel::geom_text_repel(
             data=d1,
             aes(x=year, y=data, colour=pentad, label=yr),
             inherit.aes=FALSE) +
  expand_limits(y=0) +
  scale_colour_manual(values=myColors) +
  scale_x_continuous(breaks = int_breaks) +
  labs(x="",y="", title=my_var) +
  # facet_wrap(~pentad, scales="free")
  facet_grid(stockkeylabelold~pentad, scales="free")

```

\newpage

**Pelagic stocks**

```{r, echo=FALSE, fig.asp=1.2, fig.align="center", message=FALSE, warning=FALSE}

## Pelagic

# advice
d1 <-
  iAdvice %>%
  filter(stockkeylabelold %in% pelstocks) %>%
  filter(adviceonstock) %>%
  filter(tacyear >= fy, tacyear <= ly) %>%
  dplyr::select(stockkeylabelold, year=tacyear, 
                f=fset, ssb=ssbset) %>%
  mutate(decade = as.character(10 * floor(year/10))) %>% 
  mutate(yr = stringr::str_pad(ifelse(year >=2000, year-2000, year-1900), 
                               width=2, pad="0")) %>% 
  tidyr::pivot_longer(names_to = "variable", values_to = "data", ssb:f) %>% 
  mutate(variable = factor(variable, levels=c("ssb","f","rec")))

# assess
d2 <-
  iAssess %>%
  filter(stockkeylabelold %in% pelstocks) %>%
  filter(purpose=="advice") %>% 
  # filter(assessmentyear == max(assessmentyear)) %>%
  filter(year >= fy, year <= ly) %>%
  dplyr::select(stockkeylabelold, assessmentyear, year, 
                ssb=stocksize, f=fishingpressure, rec=recruitment) %>%
  mutate(decade = as.character(10 * floor(assessmentyear/10))) %>% 
  tidyr::pivot_longer(names_to = "variable", values_to = "data", ssb:rec) %>% 
  mutate(variable = factor(variable, levels=c("ssb","f","rec")))

i <- 1
for (i in 1:length(pelstocks)) {

  assign(paste0("p",i),
         d2 %>% 
           filter(stockkeylabelold == pelstocks[[i]]) %>%

           ggplot(aes(x=year, y=data, group=assessmentyear)) +
           theme_publication() +
           geom_line(aes(colour=decade)) +
           geom_point(data=filter(d1, stockkeylabelold == pelstocks[[i]]), 
                      aes(x=year, y=data, colour=decade),
                      inherit.aes=FALSE) +
           ggrepel::geom_text_repel(
                      data=filter(d1, stockkeylabelold == pelstocks[[i]]),
                      aes(x=year, y=data, colour=decade, label=yr),
                      inherit.aes=FALSE) +
           scale_colour_manual(values=myColors) +
           expand_limits(y=0) +
           labs(x="",y="", title=pelstocks[[i]]) +
           facet_wrap(~variable, scales="free_y")
         )
}

p1 + p2 + p3 + p4 + plot_layout(ncol = 1, guides="collect") &
  theme(legend.position='bottom')

```

\newpage

**Demersal F**

```{r, echo=FALSE, fig.asp=1.4, fig.align="center", message=FALSE, warning=FALSE}

# plot of F against Fset and SSB against SSBset for multiple assessments

mytype    <- "Demersal"
myvar     <- "f"

d1 <-
  iAdvice %>%
  filter(stockkeylabelold %in% demstocks) %>%
  filter(adviceonstock) %>%
  filter(tacyear >= fay, tacyear <= lay) %>%
  dplyr::select(stockkeylabelold, year=tacyear, f=fset, ssb=ssbset) %>%
  tidyr::pivot_longer(names_to = "variable", values_to = "data", ssb:f)

d2 <-
  iAssess %>%
  filter(stockkeylabelold %in% demstocks) %>%
  filter(purpose=="advice") %>% 
  filter(assessmentyear %in% assessmentyears) %>% 
  filter(year >= fay, year <= lay) %>%
  mutate(source=paste0("assess", assessmentyear)) %>% 
  dplyr::select(stockkeylabelold, year, ssb=stocksize, f=fishingpressure, source) %>%
  tidyr::pivot_longer(names_to = "variable", values_to = "data", ssb:f)

# iAssess %>% filter(stockkeylabelold=="hke-nrtn", assessmentyear==2010) %>% View()

t <-
  d2 %>% 
  left_join(d1, by=c("stockkeylabelold", "year", "variable")) %>% 
  rename(
    assess=data.x,
    forecast=data.y
  ) %>% 
  drop_na(forecast) %>% 
  mutate(decade = as.character(10 * floor(year / 10))) %>% 
  mutate(yr = stringr::str_pad(ifelse(year >=2000, year-2000, year-1900), width=2, pad="0"))

# maxssb<- t %>% filter(variable=="ssb") %>% ungroup() %>% 
#   summarise(maxssb=max(assess, forecast, na.rm=TRUE)) %>% 
#   as.numeric()

i <- 1
for (i in 1:length(demstocks)) {

  max  <- t %>% 
    filter(variable==myvar) %>% 
    filter(stockkeylabelold == demstocks[[i]]) %>% 
    ungroup() %>%
    summarise(max=max(assess, forecast, na.rm=TRUE)) %>%
    as.numeric()

  assign(paste0("p",i),
         t %>% 
           filter(variable == myvar) %>% 
           filter(stockkeylabelold == demstocks[[i]]) %>%

           ggplot(aes(x=assess, y=forecast)) +
           theme_publication() +
           geom_point(aes(colour=decade)) +
           ggrepel::geom_text_repel(aes(colour=decade, label=yr)) +
           scale_colour_manual(values=myColors) +
           geom_abline(slope=1, intercept=0, linetype="dashed", linewidth=0.5) +
           scale_x_continuous(limits=c(0,max)) +
           scale_y_continuous(limits=c(0,max)) +
           facet_grid(stockkeylabelold~source)
         )
}

p1 + p2 + p3 + p4 + p5 + plot_layout(ncol = 1, guides="collect") &
  theme(legend.position='bottom')

```

\newpage

**Demersal SSB**

```{r, echo=FALSE, fig.asp=1.4, fig.align="center", message=FALSE, warning=FALSE}

# plot of F against Fset and SSB against SSBset for multiple assessments

mytype    <- "Demersal"
myvar     <- "ssb"

d1 <-
  iAdvice %>%
  filter(stockkeylabelold %in% demstocks) %>%
  filter(adviceonstock) %>%
  filter(tacyear >= fay, tacyear <= lay) %>%
  dplyr::select(stockkeylabelold, year=tacyear, f=fset, ssb=ssbset) %>%
  tidyr::pivot_longer(names_to = "variable", values_to = "data", ssb:f)

d2 <-
  iAssess %>%
  filter(stockkeylabelold %in% demstocks) %>%
  filter(purpose=="advice") %>% 
  filter(assessmentyear %in% assessmentyears) %>% 
  filter(year >= fay, year <= lay) %>%
  mutate(source=paste0("assess", assessmentyear)) %>% 
  dplyr::select(stockkeylabelold, year, ssb=stocksize, f=fishingpressure, source) %>%
  tidyr::pivot_longer(names_to = "variable", values_to = "data", ssb:f)

# d2 %>% filter(stockkeylabelold=="hke-nrtn", source=="assess1990") %>% View()

t <-
  d2 %>% 
  left_join(d1, by=c("stockkeylabelold", "year", "variable")) %>% 
  rename(
    assess=data.x,
    forecast=data.y
  ) %>% 
  drop_na(forecast) %>% 
  mutate(decade = as.character(10 * floor(year / 10))) %>% 
  mutate(yr = stringr::str_pad(ifelse(year >=2000, year-2000, year-1900), width=2, pad="0"))

# maxssb<- t %>% filter(variable=="ssb") %>% ungroup() %>% 
#   summarise(maxssb=max(assess, forecast, na.rm=TRUE)) %>% 
#   as.numeric()

i <- 1
for (i in 1:length(demstocks)) {

  max  <- t %>% 
    filter(variable==myvar) %>% 
    filter(stockkeylabelold == demstocks[[i]]) %>% 
    ungroup() %>%
    summarise(max=max(assess, forecast, na.rm=TRUE)) %>%
    as.numeric()

  assign(paste0("p",i),
         t %>% 
           filter(variable == myvar) %>% 
           filter(stockkeylabelold == demstocks[[i]]) %>%

           ggplot(aes(x=assess, y=forecast)) +
           theme_publication() +
           geom_point(aes(colour=decade)) +
           ggrepel::geom_text_repel(aes(colour=decade, label=yr)) +
           scale_colour_manual(values=myColors) +
           geom_abline(slope=1, intercept=0, linetype="dashed", linewidth=0.5) +
           scale_x_continuous(limits=c(0,max)) +
           scale_y_continuous(limits=c(0,max)) +
           facet_grid(stockkeylabelold~source)
         )
}

p1 + p2 + p3 + p4 + p5 + plot_layout(ncol = 1, guides="collect") &
  theme(legend.position='bottom')

```

\newpage

**Pelagic F**

```{r, echo=FALSE, fig.asp=1.2, fig.align="center", message=FALSE, warning=FALSE}

# plot of F against Fset and SSB against SSBset for multiple assessments

mytype    <- "Pelagic"
myvar     <- "f"

d1 <-
  iAdvice %>%
  filter(stockkeylabelold %in% pelstocks) %>%
  filter(adviceonstock) %>%
  filter(tacyear >= fay, tacyear <= lay) %>%
  dplyr::select(stockkeylabelold, year=tacyear, f=fset, ssb=ssbset) %>%
  tidyr::pivot_longer(names_to = "variable", values_to = "data", ssb:f)

d2 <-
  iAssess %>%
  filter(stockkeylabelold %in% pelstocks) %>%
  filter(purpose=="advice") %>% 
  filter(assessmentyear %in% assessmentyears) %>% 
  filter(year >= fay, year <= lay) %>%
  mutate(source=paste0("assess", assessmentyear)) %>% 
  dplyr::select(stockkeylabelold, year, ssb=stocksize, f=fishingpressure, source) %>%
  tidyr::pivot_longer(names_to = "variable", values_to = "data", ssb:f)

t <-
  d2 %>% 
  left_join(d1, by=c("stockkeylabelold", "year", "variable")) %>% 
  rename(
    assess=data.x,
    forecast=data.y
  ) %>% 
  drop_na(forecast) %>% 
  mutate(decade = as.character(10 * floor(year / 10))) %>% 
  mutate(yr = stringr::str_pad(ifelse(year >=2000, year-2000, year-1900), width=2, pad="0"))

# maxssb<- t %>% filter(variable=="ssb") %>% ungroup() %>% 
#   summarise(maxssb=max(assess, forecast, na.rm=TRUE)) %>% 
#   as.numeric()

i <- 1
for (i in 1:length(pelstocks)) {

  max  <- t %>% 
    filter(variable==myvar) %>% 
    filter(stockkeylabelold == pelstocks[[i]]) %>% 
    ungroup() %>%
    summarise(max=max(assess, forecast, na.rm=TRUE)) %>%
    as.numeric()

  assign(paste0("p",i),
         t %>% 
           filter(variable == myvar) %>% 
           filter(stockkeylabelold == pelstocks[[i]]) %>%

           ggplot(aes(x=assess, y=forecast)) +
           theme_publication() +
           geom_point(aes(colour=decade)) +
           ggrepel::geom_text_repel(aes(colour=decade, label=yr)) +
           scale_colour_manual(values=myColors) +
           geom_abline(slope=1, intercept=0, linetype="dashed", linewidth=0.5) +
           scale_x_continuous(limits=c(0,max)) +
           scale_y_continuous(limits=c(0,max)) +
           facet_grid(stockkeylabelold~source)
         )
}

p1 + p2 + p3 + p4 + plot_layout(ncol = 1, guides="collect") &
  theme(legend.position='bottom')

```

\newpage

**Pelagic SSB**

```{r, echo=FALSE, fig.asp=1.2, fig.align="center", message=FALSE, warning=FALSE}

# plot of F against Fset and SSB against SSBset for multiple assessments

mytype    <- "Pelagic"
myvar     <- "ssb"

d1 <-
  iAdvice %>%
  filter(stockkeylabelold %in% pelstocks) %>%
  filter(adviceonstock) %>%
  filter(tacyear >= fay, tacyear <= lay) %>%
  dplyr::select(stockkeylabelold, year=tacyear, f=fset, ssb=ssbset) %>%
  tidyr::pivot_longer(names_to = "variable", values_to = "data", ssb:f)

d2 <-
  iAssess %>%
  filter(stockkeylabelold %in% pelstocks) %>%
  filter(purpose=="advice") %>% 
  filter(assessmentyear %in% assessmentyears) %>% 
  filter(year >= fay, year <= lay) %>%
  mutate(source=paste0("assess", assessmentyear)) %>% 
  dplyr::select(stockkeylabelold, year, ssb=stocksize, f=fishingpressure, source) %>%
  tidyr::pivot_longer(names_to = "variable", values_to = "data", ssb:f)

t <-
  d2 %>% 
  left_join(d1, by=c("stockkeylabelold", "year", "variable")) %>% 
  rename(
    assess=data.x,
    forecast=data.y
  ) %>% 
  drop_na(forecast) %>% 
  mutate(decade = as.character(10 * floor(year / 10))) %>% 
  mutate(yr = stringr::str_pad(ifelse(year >=2000, year-2000, year-1900), width=2, pad="0"))

# maxssb<- t %>% filter(variable=="ssb") %>% ungroup() %>% 
#   summarise(maxssb=max(assess, forecast, na.rm=TRUE)) %>% 
#   as.numeric()

i <- 1
for (i in 1:length(pelstocks)) {

  max  <- t %>% 
    filter(variable==myvar) %>% 
    filter(stockkeylabelold == pelstocks[[i]]) %>% 
    ungroup() %>%
    summarise(max=max(assess, forecast, na.rm=TRUE)) %>%
    as.numeric()

  assign(paste0("p",i),
         t %>% 
           filter(variable == myvar) %>% 
           filter(stockkeylabelold == pelstocks[[i]]) %>%

           ggplot(aes(x=assess, y=forecast)) +
           theme_publication() +
           geom_point(aes(colour=decade)) +
           ggrepel::geom_text_repel(aes(colour=decade, label=yr)) +
           scale_colour_manual(values=myColors) +
           geom_abline(slope=1, intercept=0, linetype="dashed", linewidth=0.5) +
           scale_x_continuous(limits=c(0,max)) +
           scale_y_continuous(limits=c(0,max)) +
           facet_grid(stockkeylabelold~source)
         )
}

p1 + p2 + p3 + p4 + plot_layout(ncol = 1, guides="collect") &
  theme(legend.position='bottom')

```